---
title: "P8105 Homework 5"
author: "Shehzrin Shah"
date: "2024-11-07"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1 

Create a function for size n that randomly draws birthdays for each person, checks whether there are duplicate birthdays in the group, and returns `TRUE` or `FALSE` based on the result. 

```{r}
bday_sim = function(n) {
  bdays = sample(1:365, size = n, replace = TRUE)
  duplicate = length(unique(bdays)) < n 
  return(duplicate)
}
```

Run this function 10,000 times for each group size between 2 and 50.

```{r}
sim_res =
  expand_grid(n = 2:50,
              iter = 1:10000) |>
  mutate(res = map_lgl(n, bday_sim)) |>
  group_by(n) |>
  summarize(prob = mean(res))
```

Make a plot showing the probability that at least two people in the group will share a birthday (by averaging across 10,000 simulation runs for each group size between 2 and 50) as a function of group size. 

```{r}
sim_res |>
  ggplot(aes(x = n, y = prob)) +
  geom_line() +
  labs(title = "Probability of Shared Birthday in a Group",
       x = "Group Size",
       y = "Probability of Shared Birthday")
```

As n (# of people in a room) increases, the probability of a shared birthday increases. When the group size is about 22 people, the probability of a shared birthday is about 50%. With a group size of 50, the probability is nearly 100% which suggests it is almost certain or extremely likely that at least two people in the group will share a birthday. I expect that increasing the group size even more would, at some point, result in a 100% probability of a shared birthday. 

